Supplementary Material for Food-web complexity alters the fitness landscape of an insect herbivore

### 2018-06-07 supp_mat_complexity_selection.Rmd
some early explorations of directly calculating derivatives

## CREATE SEPARATE DATASETS FOR SELECTION ANALYSES ----
control_df <- gall_selection.df %>%
  filter(Treatment.focus == "Control") %>%
  mutate(Gall_Height_mm = Gall_Height_mm - mean(Gall_Height_mm),
         gall_individuals = gall_individuals - mean(gall_individuals),
         Density_per_100_shoots = Density_per_100_shoots - mean(Density_per_100_shoots))

treatment_df <- gall_selection.df %>%
  filter(Treatment.focus == "Ectoparasitoid exclusion") %>%
  mutate(Gall_Height_mm = Gall_Height_mm - mean(Gall_Height_mm),
         gall_individuals = gall_individuals - mean(gall_individuals),
         Density_per_100_shoots = Density_per_100_shoots - mean(Density_per_100_shoots))
```

```{r FUNCTIONS}
gamm.model <- function(formula, data) {
  gamm4(formula, data, 
        random = ~(1|Genotype/Plant_Position/Gall_Number), 
        family = binomial(link = logit))
}

gam.plot <- function(gam.model, ...) {
  plot(gam.model, seWithMean = T, shift = mean(predict(gam.model)), trans = function(x) {exp(x)/(1+exp(x))}, ...)
}

@@

get_insect_FL.df <- function(FL){
  data.frame(mean_fitness = FL$Wbar, FL$points, t(FL$WbarPI)) %>%
    mutate(lower_2.5 = X2.5., upper_97.5 = X97.5.)
}

overdisp_fun <- function(model) {
  ## number of variance parameters in 
  ##   an n-by-n variance-covariance matrix
  vpars <- function(m) {
    nrow(m)*(nrow(m)+1)/2
  }
  model.df <- sum(sapply(VarCorr(model),vpars))+length(fixef(model))
  rdf <- nrow(model.frame(model))-model.df
  rp <- residuals(model,type="pearson")
  Pearson.chisq <- sum(rp^2)
  prat <- Pearson.chisq/rdf
  pval <- pchisq(Pearson.chisq, df=rdf, lower.tail=FALSE)
  c(chisq=Pearson.chisq,ratio=prat,rdf=rdf,p=pval)
}
```

```{r}
# convert treatment to factor to use as 'by' variable in gamm.model
gall_selection.df <- gall_selection.df %>%
  mutate(Treatment_contr.sum=factor(Treatment.focus), 
         sc.gall_size = as.numeric(scale(Gall_Height_mm)),
         sc.clutch_size = as.numeric(scale(gall_individuals)),
         sc.female_preference = as.numeric(scale(Density_per_100_shoots)))
contrasts(gall_selection.df$Treatment_contr.sum) <- 'contr.sum'
# Inspect contrasts:
contrasts(gall_selection.df$Treatment_contr.sum)

## AIC MODEL SELECTION
# Note that I use gamm4 to fit all models to maintain consistency

## GAMM full
m1 <- gamm.model(gall_survival ~ Treatment_contr.sum +
                   s(sc.gall_size, by=Treatment_contr.sum) + 
                   s(sc.clutch_size, by=Treatment_contr.sum) + 
                   s(sc.female_preference, by=Treatment_contr.sum) + 
                   sc.gall_size:sc.clutch_size + sc.gall_size:sc.clutch_size:Treatment_contr.sum +
                   sc.gall_size:sc.female_preference + sc.gall_size:sc.female_preference:Treatment_contr.sum +                                 sc.clutch_size:sc.female_preference + sc.clutch_size:sc.female_preference:Treatment_contr.sum,
                 data = gall_selection.df)
summary(m1$gam); summary(m1$mer)

gamm_reduced <- gamm.model(gall_survival ~ Treatment_contr.sum +
                             s(sc.gall_size) + 
                             s(sc.clutch_size, by=Treatment_contr.sum) + 
                             s(sc.female_preference, by=Treatment_contr.sum),
                 data = gall_selection.df)
summary(gamm_reduced$gam); summary(gamm_reduced$mer)

gamm_reduced2 <- gamm.model(gall_survival ~ Treatment_contr.sum +
                             sc.gall_size + 
                             sc.clutch_size + sc.clutch_size:Treatment_contr.sum + 
                             s(sc.female_preference, by=Treatment_contr.sum),
                 data = gall_selection.df)
summary(gamm_reduced2$gam); summary(gamm_reduced2$mer)

GLMM_reduced <- gamm.model(gall_survival ~ Treatment_contr.sum +
                             sc.gall_size + 
                             sc.clutch_size + sc.clutch_size:Treatment_contr.sum + 
                             sc.female_preference + sc.female_preference:Treatment_contr.sum,
                            data = gall_selection.df)
summary(GLMM_reduced$gam); summary(GLMM_reduced$mer)

AIC(m1$mer, gamm_reduced$mer, gamm_reduced2$mer, GLMM_reduced$mer)

glmer_reduced <- glmer(gall_survival ~ Treatment_contr.sum +
                             sc.gall_size + 
                             sc.clutch_size + sc.clutch_size:Treatment_contr.sum + 
                             sc.female_preference + sc.female_preference:Treatment_contr.sum +
                             (1|Genotype/Plant_Position/Gall_Number),
                       family=binomial(link=logit),
                       data = gall_selection.df, glmerControl(optimizer="bobyqa"))
summary(glmer_reduced)
overdisp_fun(glmer_reduced) # no evidence of overdispersion

v <- visreg(fit, 'Days', 'Subject', re.form=(~Days|Subject), plot=FALSE)
plot(v)
plot(v, overlay=TRUE)

test <- visreg(glmer_reduced, xvar=c("sc.gall_size"), by="Genotype", re.form=(~1|Genotype), plot=FALSE, 
       scale="response", partial=TRUE, gg=TRUE, overlay=TRUE,
       cond = list(sc.gall_size=0, sc.clutch_size=0, sc.female_preference=0))
plot(test, partial=TRUE, overlay=TRUE)
visreg(glmer_reduced, xvar=c("sc.gall_size"), scale="response", partial=TRUE, gg=TRUE, overlay=TRUE)
visreg(glmer_reduced, xvar=c("sc.clutch_size"), by="Treatment_contr.sum", scale="response", partial=TRUE, gg=TRUE, overlay=TRUE)
visreg(glmer_reduced, xvar=c("sc.female_preference"), by="Treatment_contr.sum", scale="response", partial=TRUE, gg=TRUE, overlay=TRUE)

## GLMER full
glmer_full <- gamm.model(gall_survival ~ Treatment_contr.sum*(sc.gall_size + sc.clutch_size + sc.female_preference)^2 +
                           Treatment_contr.sum*(I(sc.gall_size^2) + I(sc.clutch_size^2) + I(sc.female_preference^2)), 
                         data = gall_selection.df)
summary(glmer_full$gam); summary(glmer_full$mer)
round(vcov(glmer_full$mer), 2) # no evidence of correlations that would confound interpretations of statistical main and interaction effects.

glmer_reduced <- gamm.model(gall_survival ~ Treatment_contr.sum*(sc.gall_size + sc.clutch_size + sc.female_preference) +
                           I(sc.gall_size^2) + I(sc.female_preference^2) + sc.gall_size:sc.clutch_size, 
                         data = gall_selection.df)
summary(glmer_reduced$gam); summary(glmer_reduced$mer, correlation=TRUE)
visreg(glmer_reduced$gam, xvar="sc.gall_size", by="sc.clutch_size", scale="response")

glmer_reduced2 <- gamm.model(gall_survival ~ Treatment_contr.sum*sc.clutch_size + sc.gall_size + sc.female_preference +
                               I(sc.female_preference^2) + sc.gall_size:sc.clutch_size, 
                         data = gall_selection.df)
summary(glmer_reduced2$gam); summary(glmer_reduced2$mer)

AIC(glmer_full$mer, glmer_reduced$mer, glmer_reduced2$mer)

## LMER full
m3 <- gamm4(gall_survival ~ Treatment_contr.sum*(sc.gall_size + sc.clutch_size + sc.female_preference)^2 +
                   Treatment_contr.sum*(I(sc.gall_size^2) + I(sc.clutch_size^2) + I(sc.female_preference^2)),
            family = gaussian(link=identity),
            random = ~(1|Genotype/Plant_Position/Gall_Number),
            data = gall_selection.df)

AIC(m1$mer, m2$mer, m3$mer) # suggests that the GLMM is best

## Do we need a model with interactions between traits?
m4 <- gamm.model(gall_survival ~ Treatment_contr.sum*(sc.gall_size + sc.clutch_size + sc.female_preference) + #^2 +
                   Treatment_contr.sum*(I(sc.gall_size^2) + I(sc.clutch_size^2) + I(sc.female_preference^2)), 
                 data = gall_selection.df)
AIC(m2$mer, m4$mer) # AIC suggests not

## Do we need a model with non-linear trait effects?
m5 <- gamm.model(gall_survival ~ Treatment_contr.sum*(sc.gall_size + sc.clutch_size + sc.female_preference) + #^2 +
                   (I(sc.gall_size^2) + I(sc.clutch_size^2) + I(sc.female_preference^2)), 
                 data = gall_selection.df)
AIC(m4$mer, m5$mer) # 
summary(m5$gam)
summary(m5$mer)

m6 <- gamm.model(gall_survival ~ Treatment_contr.sum*(sc.gall_size + sc.clutch_size + sc.female_preference), 
                 data = gall_selection.df)
AIC(m5$mer, m6$mer) 
```

@@

## CONTROL PLANTS ----
gamm_red.control <- gamm.model(gall_survival ~ 
                                  s(Gall_Height_mm) + 
                                  s(gall_individuals, k=9) + 
                                  s(Density_per_100_shoots) +
                                  Gall_Height_mm:gall_individuals + 
                                  Gall_Height_mm:Density_per_100_shoots + 
                                  gall_individuals:Density_per_100_shoots,
                                data = control_df)
summary(gamm_full.control$gam)
summary(gamm_full.control$mer)

gamm_red.control <- gamm.model(gall_survival ~ 
                                  s(Gall_Height_mm) + 
                                  s(gall_individuals, k=9) + 
                                  s(Density_per_100_shoots),
                                data = control_df)
summary(gamm_red.control$gam)
summary(gamm_red.control$mer)

gamm_red.control.noG <- gamm4(gall_survival ~ 
                                  s(Gall_Height_mm) + 
                                  s(gall_individuals, k=9) + 
                                  s(Density_per_100_shoots),
                                  random = ~(1|Plant_Position/Gall_Number), 
                                  family = binomial(link = logit),
                                data = control_df)
summary(gamm_red.control.noG$gam)
summary(gamm_red.control.noG$mer)

glmm_full.control <- gamm.model(gall_survival ~ 
                                  poly(Gall_Height_mm,2) + poly(gall_individuals,2) + poly(Density_per_100_shoots,2) +
                                  Gall_Height_mm:gall_individuals + 
                                  Gall_Height_mm:Density_per_100_shoots +
                                  gall_individuals:Density_per_100_shoots,
                                data = control_df)
summary(glmm_full.control$gam)
summary(glmm_full.control$mer)

glmm_red.control <- gamm.model(gall_survival ~ 
                                  poly(Gall_Height_mm,2) + poly(gall_individuals,2) + poly(Density_per_100_shoots,2),
                                data = control_df)
summary(glmm_red.control$gam)
summary(glmm_red.control$mer)

lmm_red.control <- gamm4(gall_survival ~ 
                                  poly(Gall_Height_mm,2) + poly(gall_individuals,2) + poly(Density_per_100_shoots,2),
                           random = ~(1|Plant_Position/Gall_Number),
                                data = control_df, family=gaussian())
summary(lmm_red.control$gam)
summary(lmm_red.control$mer)

AIC(gamm_full.control$mer, gamm_red.control$mer, gamm_red.control.noG$mer, glmm_full.control$mer, glmm_red.control$mer, lmm_red.control$mer)

## TREATMENT PLANTS ----
gamm_red.treatment <- gamm.model(gall_survival ~ 
                                  s(Gall_Height_mm) + 
                                  s(gall_individuals, k=9) + 
                                  s(Density_per_100_shoots) +
                                  Gall_Height_mm:gall_individuals + 
                                  Gall_Height_mm:Density_per_100_shoots + 
                                  gall_individuals:Density_per_100_shoots,
                                data = treatment_df)
summary(gamm_full.treatment$gam)
summary(gamm_full.treatment$mer)

gamm_red.treatment <- gamm.model(gall_survival ~ 
                                  s(Gall_Height_mm) + 
                                  s(gall_individuals, k=9) + 
                                  s(Density_per_100_shoots),
                                data = treatment_df)
summary(gamm_red.treatment$gam)
summary(gamm_red.treatment$mer)

gamm_red.treatment.noG <- gamm4(gall_survival ~ 
                                  s(Gall_Height_mm) + 
                                  s(gall_individuals, k=9) + 
                                  s(Density_per_100_shoots),
                                  random = ~(1|Plant_Position/Gall_Number), 
                                  family = binomial(link = logit),
                                data = treatment_df)
summary(gamm_red.treatment.noG$gam)
summary(gamm_red.treatment.noG$mer)

glmm_full.treatment <- gamm.model(gall_survival ~ 
                                  poly(Gall_Height_mm,2) + poly(gall_individuals,2) + poly(Density_per_100_shoots,2) +
                                  Gall_Height_mm:gall_individuals + 
                                  Gall_Height_mm:Density_per_100_shoots +
                                  gall_individuals:Density_per_100_shoots,
                                data = treatment_df)
summary(glmm_full.treatment$gam)
summary(glmm_full.treatment$mer)

glmm_red.treatment <- gamm.model(gall_survival ~ 
                                  poly(Gall_Height_mm,2) + poly(gall_individuals,2) + poly(Density_per_100_shoots,2),
                                data = treatment_df)
summary(glmm_red.treatment$gam)
summary(glmm_red.treatment$mer)

lmm_red.treatment <- gamm4(gall_survival ~ 
                                  poly(Gall_Height_mm,2) + poly(gall_individuals,2) + poly(Density_per_100_shoots,2),
                           random = ~(1|Plant_Position/Gall_Number),
                                data = treatment_df, family=gaussian())
summary(lmm_red.treatment$gam)
summary(lmm_red.treatment$mer)

AIC(gamm_full.treatment$mer, gamm_red.treatment$mer, gamm_red.treatment.noG$mer, glmm_full.treatment$mer, glmm_red.treatment$mer, lmm_red.treatment$mer)

## Linear mixed-models
lmm_full.control <- lmer(gall_survival/0.45 ~ 
                                  poly(Gall_Height_mm,2) + poly(gall_individuals,2) + poly(Density_per_100_shoots,2) +
                                  Gall_Height_mm:gall_individuals + 
                                  Gall_Height_mm:Density_per_100_shoots +
                                  gall_individuals:Density_per_100_shoots +
                           (1|Genotype/Plant_Position/Gall_Number),
                                data = control_df)
summary(lmm_full.control)
car::Anova(lmm_full.control, test="F")
visreg::visreg(lmm_full.control)

lmm_full.treatment <- lmer(gall_survival/0.6 ~ 
                                  poly(Gall_Height_mm,2) + poly(gall_individuals,2) + poly(Density_per_100_shoots,2) +
                                  Gall_Height_mm:gall_individuals + 
                                  Gall_Height_mm:Density_per_100_shoots +
                                  gall_individuals:Density_per_100_shoots +
                             (1|Genotype/Plant_Position/Gall_Number),
                                data = treatment_df)
summary(lmm_full.treatment)
car::Anova(lmm_full.treatment, test="F")

lmm_red.treatment <- lmer(gall_survival/0.6 ~ 
                                  poly(Gall_Height_mm,2) + poly(gall_individuals,2) + poly(Density_per_100_shoots,2) +
                              (1|Genotype/Plant_Position/Gall_Number),
                                data = treatment_df)
summary(lmm_red.treatment)
car::Anova(lmm_red.treatment, test="F")
visreg::visreg(lmm_red.treatment)

AIC(lmm_full.treatment, lmm_red.treatment)

gall_selection.df$relative.fit <- with(gall_selection.df, ifelse(Treatment.focus=="Control", gall_survival/0.45, gall_survival/0.6))

lmm_full <- lmer(relative.fit ~ poly(Gall_Height_mm,2) + poly(gall_individuals,2) + poly(Density_per_100_shoots,2) + 
                                poly(Gall_Height_mm,2):Treatment.focus + 
                                poly(gall_individuals,2):Treatment.focus + 
                                poly(Density_per_100_shoots,2):Treatment.focus +
                                (1|Genotype/Plant_Position/Gall_Number),
                                data = gall_selection.df)
summary(lmm_full)
car::Anova(lmm_full, test="F")
visreg::visreg(lmm_full, xvar="Gall_Height_mm", by="Treatment.focus")
visreg::visreg(lmm_full, xvar="gall_individuals", by="Treatment.focus")
visreg::visreg(lmm_full, xvar="Density_per_100_shoots", by="Treatment.focus")
```
Parametric bootstrapping: Currently reported results

>>>>

## CREATE SEPARATE DATASETS FOR SELECTION ANALYSES ----
control_df <- gall_selection.df %>%
  filter(Treatment.focus == "Control") #%>%
  #mutate(Gall_Height_mm = Gall_Height_mm - mean(Gall_Height_mm),
  #       gall_individuals = gall_individuals - mean(gall_individuals),
  #       Density_per_100_shoots = Density_per_100_shoots - mean(Density_per_100_shoots))

treatment_df <- gall_selection.df %>%
  filter(Treatment.focus == "Ectoparasitoid exclusion") #%>%
  #mutate(Gall_Height_mm = Gall_Height_mm - mean(Gall_Height_mm),
  #       gall_individuals = gall_individuals - mean(gall_individuals),
  #       Density_per_100_shoots = Density_per_100_shoots - mean(Density_per_100_shoots))
```

Custom functions to avoid repetitive code
```{r FUNCTIONS}
gamm.model <- function(formula, data) {
  gamm4(formula, data, 
        random = ~(1|Genotype/Plant_Position/Gall_Number), 
        family = binomial(link = logit))
}

gam.plot <- function(gam.model, ...) {
  plot(gam.model, se = TRUE, seWithMean = T, shift = mean(predict(gam.model)), trans = function(x) {exp(x)/(1+exp(x))}, ...)
}

@@

get_insect_FL.df <- function(FL){
  data.frame(mean_fitness = FL$Wbar, FL$points, t(FL$WbarPI)) %>%
    mutate(lower_2.5 = X2.5., upper_97.5 = X97.5.)
}
```

@@

## CONTROL PLANTS ----
gamm_full.control <- gamm.model(gall_survival ~ s(Gall_Height_mm) + s(gall_individuals, k=9) + s(Density_per_100_shoots) +
                                  Gall_Height_mm:gall_individuals + Gall_Height_mm:Density_per_100_shoots + gall_individuals:Density_per_100_shoots,
                                data = control_df)
summary(gamm_full.control$gam); summary(gamm_full.control$mer)

glmm_full.control <- gamm.model(gall_survival ~ Gall_Height_mm + gall_individuals + Density_per_100_shoots +
                                  I(Gall_Height_mm^2) + I(gall_individuals^2) + I(Density_per_100_shoots^2) +
                                  Gall_Height_mm:gall_individuals + Gall_Height_mm:Density_per_100_shoots + gall_individuals:Density_per_100_shoots,
                                data = control_df)
summary(glmm_full.control$gam); summary(glmm_full.control$mer)

gamm_red.control <- gamm.model(gall_survival ~ s(Gall_Height_mm) + s(gall_individuals, k=9) + s(Density_per_100_shoots),
                                data = control_df)
summary(gamm_red.control$gam); summary(gamm_red.control$mer)

glmm_quad.control <- gamm.model(gall_survival ~ Gall_Height_mm + gall_individuals + Density_per_100_shoots +
                                  I(Gall_Height_mm^2) + I(gall_individuals^2) + I(Density_per_100_shoots^2),
                                data = control_df)
summary(glmm_quad.control$gam); summary(glmm_quad.control$mer)

glmm_lin.control <- gamm.model(gall_survival ~ Gall_Height_mm + gall_individuals + Density_per_100_shoots,
                                data = control_df)
summary(glmm_lin.control$gam); summary(glmm_lin.control$mer)

lmm_lin.control <- gamm4(gall_survival ~ Gall_Height_mm + gall_individuals + Density_per_100_shoots,
                          data = control_df,
                          random = ~(1|Genotype/Plant_Position/Gall_Number), 
                          family = gaussian(link = identity))
summary(lmm_lin.control$gam); summary(lmm_lin.control$mer)

glmm_lin.red.control <- gamm.model(gall_survival ~ Gall_Height_mm,
                                data = control_df)
summary(glmm_lin.red.control$gam); summary(glmm_lin.red.control$mer)

glmm_null.control <- gamm.model(gall_survival ~ 1,
                                data = control_df)
summary(glmm_null.control$gam); summary(glmm_null.control$mer)

glmm_lin.red.noRE.control <- gamm4(gall_survival ~ Gall_Height_mm,
                          data = control_df,
                          random = ~(1|Gall_Number), 
                          family = binomial(link = logit))
summary(glmm_lin.red.noRE.control$gam); summary(glmm_lin.red.noRE.control$mer)

AIC(gamm_full.control$mer, glmm_full.control$mer, gamm_red.control$mer, glmm_quad.control$mer, glmm_lin.control$mer, lmm_lin.control$mer, glmm_lin.red.control$mer, glmm_null.control$mer, glmm_lin.red.noRE.control$mer)

## TREATMENT PLANTS ----
gamm_full.treatment <- gamm.model(gall_survival ~ s(Gall_Height_mm) + s(gall_individuals, k=9) + s(Density_per_100_shoots) +
                                  Gall_Height_mm:gall_individuals + Gall_Height_mm:Density_per_100_shoots + gall_individuals:Density_per_100_shoots,
                                data = treatment_df)
summary(gamm_full.treatment$gam); summary(gamm_full.treatment$mer)

glmm_full.treatment <- gamm.model(gall_survival ~ Gall_Height_mm + gall_individuals + Density_per_100_shoots +
                                  I(Gall_Height_mm^2) + I(gall_individuals^2) + I(Density_per_100_shoots^2) +
                                  Gall_Height_mm:gall_individuals + Gall_Height_mm:Density_per_100_shoots + gall_individuals:Density_per_100_shoots,
                                data = treatment_df)
summary(glmm_full.treatment$gam); summary(glmm_full.treatment$mer)

gamm_red.treatment <- gamm.model(gall_survival ~ s(Gall_Height_mm) + s(gall_individuals, k=9) + s(Density_per_100_shoots),
                                data = treatment_df)
summary(gamm_red.treatment$gam); summary(gamm_red.treatment$mer)

glmm_quad.treatment <- gamm.model(gall_survival ~ Gall_Height_mm + gall_individuals + Density_per_100_shoots +
                                  I(Gall_Height_mm^2) + I(gall_individuals^2) + I(Density_per_100_shoots^2),
                                data = treatment_df)
summary(glmm_quad.treatment$gam); summary(glmm_quad.treatment$mer)

glmm_lin.treatment <- gamm.model(gall_survival ~ Gall_Height_mm + gall_individuals + Density_per_100_shoots,
                                data = treatment_df)
summary(glmm_lin.treatment$gam); summary(glmm_lin.treatment$mer)

lmm_lin.treatment <- gamm4(gall_survival ~ Gall_Height_mm + gall_individuals + Density_per_100_shoots,
                          data = treatment_df,
                          random = ~(1|Genotype/Plant_Position/Gall_Number), 
                          family = gaussian(link = identity))
summary(lmm_lin.treatment$gam); summary(lmm_lin.treatment$mer)

glmm_quad.red.treatment <- gamm.model(gall_survival ~ Gall_Height_mm + gall_individuals + Density_per_100_shoots + I(Density_per_100_shoots^2),
                                data = treatment_df)
summary(glmm_quad.red.treatment$gam); summary(glmm_quad.red.treatment$mer)

glmm_null.treatment <- gamm.model(gall_survival ~ 1,
                                data = treatment_df)
summary(glmm_null.treatment$gam); summary(glmm_null.treatment$mer)

glmm_lin.red.noRE.treatment <- gamm4(gall_survival ~ Gall_Height_mm,
                          data = treatment_df,
                          random = ~(1|Gall_Number), 
                          family = binomial(link = logit))
summary(glmm_lin.red.noRE.treatment$gam); summary(glmm_lin.red.noRE.treatment$mer)

AIC(gamm_full.treatment$mer, glmm_full.treatment$mer, gamm_red.treatment$mer, glmm_quad.treatment$mer, glmm_lin.treatment$mer, lmm_lin.treatment$mer, glmm_lin.red.treatment$mer, glmm_null.treatment$mer, glmm_lin.red.noRE.treatment$mer, glmm_quad.red.treatment$mer)
```

```{r}
library(schoenberg)
pointwise <- confint(gamm_red.control$gam, parm = "Gall_Height_mm", trans = function(x) {exp(x)/(1+exp(x))})
simultaneous <- confint(gamm_red.control$gam, parm = "Gall_Height_mm", trans = function(x) {exp(x)/(1+exp(x))}, type = "simultaneous", seWithMean=TRUE)

compare.confint <- bind_rows(mutate(pointwise, confint = "pointwise"), mutate(simultaneous, confint = "simultaneous")) 
compare.confint$confint <- relevel(factor(compare.confint$confint), ref = "simultaneous")

ggplot(compare.confint, aes(x=Gall_Height_mm)) +
  geom_ribbon(aes(ymin=lower, ymax=upper), alpha=0.25) +
  geom_line(aes(y=est), color="black") +
  facet_wrap(~confint)

schluter <- gam.plot(gamm_red.control$gam)
schluter

gam.plot(glmm_lin.red.control$gam, all.terms=TRUE)
test <- plot(glmm_lin.red.control$gam, all.terms=TRUE)
gam.gradients(mod = glmm_lin.red.control$gam, phenotype = "Gall_Height_mm", se.method = "n")
coef(glmm_lin.red.control$gam)

# test derivative calculations
control_df$Genotype <- factor(control_df$Genotype)
control_df$Plant_Position <- factor(control_df$Plant_Position)
control_df$Gall_Number <- factor(control_df$Gall_Number)
A <- gam(gall_survival ~ s(Gall_Height_mm) + s(Plant_Position, bs="re") + s(Gall_Number, bs="re"), data = control_df, family=binomial(link=logit)) # random=~(1|Genotype/Plant_Position/Gall_Number),
summary(A)

head(control_df)
A_withMean <- plot(A, se = TRUE, seWithMean = T, shift = mean(predict(A)), trans = function(x) {exp(x)/(1+exp(x))}); A_withMean
A_only <- plot(A, se = TRUE, shift = mean(predict(A)), trans = function(x) {exp(x)/(1+exp(x))}); A_only
draw(A)

# new data for prediction
newDF <- with(control_df, data.frame(Gall_Height_mm = seq(min(Gall_Height_mm),max(Gall_Height_mm),length.out = 100)))
# prediction of smoothed estimates at each unique year value
# with standard error    
B <- predict(A$gam,  newDF, type="response", se.fit=TRUE)
B

# finite difference approach to derivatives following
# example from ?predict.gam

eps <- 1e-7
X0 <- predict(A, newDF, type = 'lpmatrix')
X0 <- exp(X0)/(1+exp(X0))

newDFeps_p <- newDF + eps

X1 <- predict(A, newDFeps_p, type = 'lpmatrix')
X1 <- exp(X1)/(1+exp(X1))
# finite difference approximation of first derivative
# the design matrix
Xp <- (X0 - X1) / eps
Xp
# first derivative
fd_d1 <- Xp %*% coef(A)
fd_d1/mean_fit
# second derivative
newDFeps_m <- newDF - eps

X_1 <- predict(A$gam, newDFeps_m, type = 'lpmatrix')
# design matrix for second derivative
Xpp <- (X1 + X_1 - 2*X0)  / eps^2
# second derivative
fd_d2 <- Xpp %*% coef(A$gam)

exp(fd_d1) / ( 1+exp(fd_d1))  # type = "response"
exp(fd_d2) / ( 1+exp(fd_d2))  # type = "response"

mean(fd_d1)/mean_fit
mean(fd_d2)/mean_fit

gam.gradients(A, phenotype = "Gall_Height_mm", se.method="n", standardized = FALSE)

coef(A)
mean_fit <- mean(control_df$gall_survival)
fderiv_size <- fderiv(A, newdata = data.frame(Gall_Height_mm = mean(control_df$Gall_Height_mm)), offset = mean_fit)
test <- fderiv_size$derivatives$`s(Gall_Height_mm)`$deriv
exp(test)/(1+exp(test))

plot(A, se = TRUE, seWithMean = T, shift = mean(predict(A)), trans = function(x) {exp(x)/(1+exp(x))})
diff <- .Machine$double.eps^(1/3)
predicts <- predict(A, newdata = data.frame(Gall_Height_mm = c(mean(control_df$Gall_Height_mm)-diff,mean(control_df$Gall_Height_mm)+diff)), type="lpmatrix")
tests <- predicts %*% coef(A)
exp(tests)/(1+exp(tests))
t1 <- exp(predicts[2])/(1+exp(predicts[2]));t1
t2 <- exp(predicts[1])/(1+exp(predicts[1]));t2
((t1-t2)/(diff*2))/mean(control_df$gall_survival)
gam.gradients(A, phenotype = "Gall_Height_mm", se.method="n", standardized = FALSE)

	Wbar <- function(x, mod, phenotype,covariates) {
    		new.d <- as.data.frame(mod$model[,c(phenotype,covariates)])
    		names(new.d)<-c(phenotype,covariates)
    		new.d2<-new.d
    		for (i in 1:length(x)) {
    		    new.d[,as.character(phenotype[i])] <-
    		    new.d[,as.character(phenotype[i])]+x[i]
    		}

    		p <- predict.gam(
			   object=mod,
    	 	   newdata=new.d,
    	 	   newdata.guaranteed=TRUE,
    	 	   type="response"
    		)
    	return(mean(p))
	}
	as.data.frame(A$model[ ,"Gall_Height_mm"])
	#mean(control_df$gall_survival)
	Wbar(x=0, mod=A, phenotype="Gall_Height_mm", covariates=NULL)
first.derivatives <- grad(f=Wbar, x=rep(0, 1), mod=A, phenotype="Gall_Height_mm", covariates=NULL); first.derivatives
denom <- Wbar(x=rep(0,1), mod=A, phenotype="Gall_Height_mm", covariates=NULL); denom
first.derivatives/denom
pracma:::grad
A$Vp
A$coefficients
A_withMean
```

Parametric bootstrapping: Currently reported results
